---
title: "Analyse_Gene_Clusters"
author: "Jaime Ken Costa Pereira"
date: "1/11/2021"
output: html_document
---

load R pacakges

```{r}
library(dplyr)
library("AnnotationDbi")
library("org.Hs.eg.db")
library("ALL")
library("hgu95av2.db")
library("GO.db")
library("annotate")
library("genefilter")
library("GOstats")
library("RColorBrewer")
library("xtable")
library("Rgraphviz")
library(ggplot2)
library(xtable)
```

```{r pressure, echo=FALSE}
t_sne_dbscan_df <- read.csv("gene_network_matrix_tsne_dbscan.csv", header=TRUE) 
t_sne_dbscan_df$DBSCAN_cluster <- factor(t_sne_dbscan_df$DBSCAN_cluster)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
dbscan_plot <- ggplot(data=t_sne_dbscan_df, aes(tsne_1, tsne_2, colour = t_sne_dbscan_df$DBSCAN_cluster))+ 
  geom_point() +
  scale_colour_manual(values=setNames(t_sne_dbscan_df$color_codes, t_sne_dbscan_df$DBSCAN_cluster), name= "Clusters")+
  labs(x="T-SNE component 1", y="T-SNE component 2", title = "DBSCAN Cluster Labels")

png("gene_network_matrix_tsne_dbscan.png")
print(dbscan_plot)
dev.off()
```

```{r pressure, echo=FALSE}
tsne_plot <- ggplot(data=t_sne_dbscan_df, aes(tsne_1, tsne_2)) + 
  geom_point() + labs(x="T-SNE component 1", y="T-SNE component 2", title = "T-SNE Embedding of Data")

png("gene_network_matrix_tsne.png")
print(tsne_plot)
dev.off()
```


GoSTATS uses EntreZ ID's, therefore, I convert the ENSEMBLE gene IDs to Entrez ID's. Since a few genes are not converted, these are then removed.A total of 231 genes are removed.

```{r}
gene_cluster_columns <- c("GENES", "DBSCAN_cluster")
Gene_cluster_df <- t_sne_dbscan_df[gene_cluster_columns]
gene_ensemble_IDS <- Gene_cluster_df[,c('GENES')]
rownames(Gene_cluster_df) <- gene_ensemble_IDS
Gene_cluster_df$entrez = mapIds(org.Hs.eg.db,
                    keys=row.names(Gene_cluster_df), 
                    column="ENTREZID",
                    keytype="ENSEMBL",
                    multiVals="first")

Gene_cluster_df <- na.omit(Gene_cluster_df)
Gene_cluster_df <- Gene_cluster_df[,c("entrez", "DBSCAN_cluster")]
Gene_cluster_df <- distinct(Gene_cluster_df, entrez, .keep_all = TRUE)
gene_entrez_IDS <- Gene_cluster_df$entrez
rownames(Gene_cluster_df) <- Gene_cluster_df[,c("entrez")]

```


Seperate the 23 clusters
```{r}
gene_universe <- as.vector(Gene_cluster_df$entrez)

for(i in 0:22) { 
 nam <- paste("cluster_", i, sep = "")
 assign(nam, Gene_cluster_df[Gene_cluster_df$DBSCAN_cluster == i,] %>% rownames()) 
}

```



Analysis of Cluster 0 - overrepresented and underrepresented terms

```{r}
hgCutoff <- 0.05

params <- new("GOHyperGParams", geneIds = cluster_0, universeGeneIds = gene_universe,
              annotation = "hgu95av2", ontology = "BP", pvalueCutoff = hgCutoff, conditional = FALSE,
              testDirection = "over")

#Change conditional to true to do a conditional hypogeometric test
hgOver <- hyperGTest(params)
GO_over_cluster_0 <- summary(hgOver)

write.csv(GO_over_cluster_0, "GO_over_cluster_0.csv")


#  test for under represented terms
paramsUnder <- params
testDirection(paramsUnder) <- "under"
hgUnder <- hyperGTest(paramsUnder)
Go_under_cluster_0 <- summary(hgUnder)

write.csv(Go_under_cluster_0, "Go_under_cluster_0.csv")

```

Analysis of Cluster 1

```{r}
hgCutoff <- 0.05

params <- new("GOHyperGParams", geneIds = cluster_1, universeGeneIds = gene_universe,
              annotation = "hgu95av2", ontology = "BP", pvalueCutoff = hgCutoff, conditional = FALSE,
              testDirection = "over")

#Change conditional to true to do a conditional hypogeometric test
hgOver <- hyperGTest(params)
GO_over_cluster_1 <- summary(hgOver)

write.csv(GO_over_cluster_1, "GO_over_cluster_1.csv")


#  test for under represented terms
paramsUnder <- params
testDirection(paramsUnder) <- "under"
hgUnder <- hyperGTest(paramsUnder)
Go_under_cluster_1 <- summary(hgUnder)

write.csv(Go_under_cluster_1, "Go_under_cluster_1.csv")


```


Analysis of cluster_2

```{r}
hgCutoff <- 0.05

params <- new("GOHyperGParams", geneIds = cluster_2, universeGeneIds = gene_universe,
              annotation = "hgu95av2", ontology = "BP", pvalueCutoff = hgCutoff, conditional = FALSE,
              testDirection = "over")

#Change conditional to true to do a conditional hypogeometric test
hgOver <- hyperGTest(params)
GO_over_cluster_2 <- summary(hgOver)

write.csv(GO_over_cluster_2, "GO_over_cluster_2.csv")


#  test for under represented terms
paramsUnder <- params
testDirection(paramsUnder) <- "under"
hgUnder <- hyperGTest(paramsUnder)
Go_under_cluster_2 <- summary(hgUnder)

write.csv(Go_under_cluster_2, "Go_under_cluster_2.csv")

```

Analysis of cluster_3

```{r}
hgCutoff <- 0.05

params <- new("GOHyperGParams", geneIds = cluster_3, universeGeneIds = gene_universe,
              annotation = "hgu95av2", ontology = "BP", pvalueCutoff = hgCutoff, conditional = FALSE,
              testDirection = "over")

#Change conditional to true to do a conditional hypogeometric test
hgOver <- hyperGTest(params)
GO_over_cluster_3 <- summary(hgOver)

write.csv(GO_over_cluster_3, "GO_over_cluster_3.csv")


#  test for under represented terms
paramsUnder <- params
testDirection(paramsUnder) <- "under"
hgUnder <- hyperGTest(paramsUnder)
Go_under_cluster_3 <- summary(hgUnder)

write.csv(Go_under_cluster_3, "Go_under_cluster_3.csv")
```

Analysis of cluster_4

```{r}
hgCutoff <- 0.05

params <- new("GOHyperGParams", geneIds = cluster_4, universeGeneIds = gene_universe,
              annotation = "hgu95av2", ontology = "BP", pvalueCutoff = hgCutoff, conditional = FALSE,
              testDirection = "over")

#Change conditional to true to do a conditional hypogeometric test
hgOver <- hyperGTest(params)
GO_over_cluster_4 <- summary(hgOver)

write.csv(GO_over_cluster_4, "GO_over_cluster_4.csv")


#  test for under represented terms
paramsUnder <- params
testDirection(paramsUnder) <- "under"
hgUnder <- hyperGTest(paramsUnder)
Go_under_cluster_4 <- summary(hgUnder)

write.csv(Go_under_cluster_4, "Go_under_cluster_4.csv")

```

Analysis of cluster_5

```{r}
hgCutoff <- 0.05

params <- new("GOHyperGParams", geneIds = cluster_5, universeGeneIds = gene_universe,
              annotation = "hgu95av2", ontology = "BP", pvalueCutoff = hgCutoff, conditional = FALSE,
              testDirection = "over")

#Change conditional to true to do a conditional hypogeometric test
hgOver <- hyperGTest(params)
GO_over_cluster_5 <- summary(hgOver)

write.csv(GO_over_cluster_5, "GO_over_cluster_5.csv")


#  test for under represented terms
paramsUnder <- params
testDirection(paramsUnder) <- "under"
hgUnder <- hyperGTest(paramsUnder)
Go_under_cluster_5 <- summary(hgUnder)

write.csv(Go_under_cluster_5, "Go_under_cluster_5.csv")

```

Analysis of cluster_6

```{r}
hgCutoff <- 0.05

params <- new("GOHyperGParams", geneIds = cluster_6, universeGeneIds = gene_universe,
              annotation = "hgu95av2", ontology = "BP", pvalueCutoff = hgCutoff, conditional = FALSE,
              testDirection = "over")

#Change conditional to true to do a conditional hypogeometric test
hgOver <- hyperGTest(params)
GO_over_cluster_6 <- summary(hgOver)

write.csv(GO_over_cluster_6, "GO_over_cluster_6.csv")


#  test for under represented terms
paramsUnder <- params
testDirection(paramsUnder) <- "under"
hgUnder <- hyperGTest(paramsUnder)
Go_under_cluster_6 <- summary(hgUnder)

write.csv(Go_under_cluster_6, "Go_under_cluster_6.csv")

```

Analysis of cluster_7

```{r}
hgCutoff <- 0.05
hgCutoff <- 0.05

params <- new("GOHyperGParams", geneIds = cluster_7, universeGeneIds = gene_universe,
              annotation = "hgu95av2", ontology = "BP", pvalueCutoff = hgCutoff, conditional = FALSE,
              testDirection = "over")

#Change conditional to true to do a conditional hypogeometric test
hgOver <- hyperGTest(params)
GO_over_cluster_7 <- summary(hgOver)

write.csv(GO_over_cluster_7, "GO_over_cluster_7.csv")


#  test for under represented terms
paramsUnder <- params
testDirection(paramsUnder) <- "under"
hgUnder <- hyperGTest(paramsUnder)
Go_under_cluster_7 <- summary(hgUnder)

write.csv(Go_under_cluster_7, "Go_under_cluster_7.csv")

```

Analysis of cluster_8

```{r}
hgCutoff <- 0.05

params <- new("GOHyperGParams", geneIds = cluster_8, universeGeneIds = gene_universe,
              annotation = "hgu95av2", ontology = "BP", pvalueCutoff = hgCutoff, conditional = FALSE,
              testDirection = "over")

#Change conditional to true to do a conditional hypogeometric test
hgOver <- hyperGTest(params)
GO_over_cluster_8 <- summary(hgOver)

write.csv(GO_over_cluster_8, "GO_over_cluster_8.csv")


#  test for under represented terms
paramsUnder <- params
testDirection(paramsUnder) <- "under"
hgUnder <- hyperGTest(paramsUnder)
Go_under_cluster_8 <- summary(hgUnder)

write.csv(Go_under_cluster_8, "Go_under_cluster_8.csv")

```

Analysis of cluster_9

```{r}
hgCutoff <- 0.05

params <- new("GOHyperGParams", geneIds = cluster_9, universeGeneIds = gene_universe,
              annotation = "hgu95av2", ontology = "BP", pvalueCutoff = hgCutoff, conditional = FALSE,
              testDirection = "over")

#Change conditional to true to do a conditional hypogeometric test
hgOver <- hyperGTest(params)
GO_over_cluster_9 <- summary(hgOver)

write.csv(GO_over_cluster_9, "GO_over_cluster_9.csv")


#  test for under represented terms
paramsUnder <- params
testDirection(paramsUnder) <- "under"
hgUnder <- hyperGTest(paramsUnder)
Go_under_cluster_9 <- summary(hgUnder)

write.csv(Go_under_cluster_9, "Go_under_cluster_9.csv")

```

Analysis of cluster_10

```{r}
hgCutoff <- 0.05

params <- new("GOHyperGParams", geneIds = cluster_10, universeGeneIds = gene_universe,
              annotation = "hgu95av2", ontology = "BP", pvalueCutoff = hgCutoff, conditional = FALSE,
              testDirection = "over")

#Change conditional to true to do a conditional hypogeometric test
hgOver <- hyperGTest(params)
GO_over_cluster_10 <- summary(hgOver)

write.csv(GO_over_cluster_10, "GO_over_cluster_10.csv")


#  test for under represented terms
paramsUnder <- params
testDirection(paramsUnder) <- "under"
hgUnder <- hyperGTest(paramsUnder)
Go_under_cluster_10 <- summary(hgUnder)

write.csv(Go_under_cluster_10, "Go_under_cluster_10.csv")

```

Analysis of cluster_11

```{r}
hgCutoff <- 0.05

params <- new("GOHyperGParams", geneIds = cluster_11, universeGeneIds = gene_universe,
              annotation = "hgu95av2", ontology = "BP", pvalueCutoff = hgCutoff, conditional = FALSE,
              testDirection = "over")

#Change conditional to true to do a conditional hypogeometric test
hgOver <- hyperGTest(params)
GO_over_cluster_11 <- summary(hgOver)

write.csv(GO_over_cluster_11, "GO_over_cluster_11.csv")


#  test for under represented terms
paramsUnder <- params
testDirection(paramsUnder) <- "under"
hgUnder <- hyperGTest(paramsUnder)
Go_under_cluster_11 <- summary(hgUnder)

write.csv(Go_under_cluster_11, "Go_under_cluster_11.csv")

```

Analysis of cluster_12
```{r}
hgCutoff <- 0.05

params <- new("GOHyperGParams", geneIds = cluster_12, universeGeneIds = gene_universe,
              annotation = "hgu95av2", ontology = "BP", pvalueCutoff = hgCutoff, conditional = FALSE,
              testDirection = "over")

#Change conditional to true to do a conditional hypogeometric test
hgOver <- hyperGTest(params)
GO_over_cluster_12 <- summary(hgOver)

write.csv(GO_over_cluster_12, "GO_over_cluster_12.csv")


#  test for under represented terms
paramsUnder <- params
testDirection(paramsUnder) <- "under"
hgUnder <- hyperGTest(paramsUnder)
Go_under_cluster_12 <- summary(hgUnder)

write.csv(Go_under_cluster_12, "Go_under_cluster_12.csv")

```

Analysis of cluster_13

```{r}
hgCutoff <- 0.05

params <- new("GOHyperGParams", geneIds = cluster_13, universeGeneIds = gene_universe,
              annotation = "hgu95av2", ontology = "BP", pvalueCutoff = hgCutoff, conditional = FALSE,
              testDirection = "over")

#Change conditional to true to do a conditional hypogeometric test
hgOver <- hyperGTest(params)
GO_over_cluster_13 <- summary(hgOver)

write.csv(GO_over_cluster_13, "GO_over_cluster_13.csv")


#  test for under represented terms
paramsUnder <- params
testDirection(paramsUnder) <- "under"
hgUnder <- hyperGTest(paramsUnder)
Go_under_cluster_13 <- summary(hgUnder)

write.csv(Go_under_cluster_13, "Go_under_cluster_13.csv")

```

Analysis of cluster_14

```{r}
hgCutoff <- 0.05

params <- new("GOHyperGParams", geneIds = cluster_14, universeGeneIds = gene_universe,
              annotation = "hgu95av2", ontology = "BP", pvalueCutoff = hgCutoff, conditional = FALSE,
              testDirection = "over")

#Change conditional to true to do a conditional hypogeometric test
hgOver <- hyperGTest(params)
GO_over_cluster_14 <- summary(hgOver)

write.csv(GO_over_cluster_14, "GO_over_cluster_14.csv")


#  test for under represented terms
paramsUnder <- params
testDirection(paramsUnder) <- "under"
hgUnder <- hyperGTest(paramsUnder)
Go_under_cluster_14 <- summary(hgUnder)

write.csv(Go_under_cluster_14, "Go_under_cluster_14.csv")

```

Analysis of cluster_15

```{r}
hgCutoff <- 0.05

params <- new("GOHyperGParams", geneIds = cluster_15, universeGeneIds = gene_universe,
              annotation = "hgu95av2", ontology = "BP", pvalueCutoff = hgCutoff, conditional = FALSE,
              testDirection = "over")

#Change conditional to true to do a conditional hypogeometric test
hgOver <- hyperGTest(params)
GO_over_cluster_15 <- summary(hgOver)

write.csv(GO_over_cluster_15, "GO_over_cluster_15.csv")


#  test for under represented terms
paramsUnder <- params
testDirection(paramsUnder) <- "under"
hgUnder <- hyperGTest(paramsUnder)
Go_under_cluster_15 <- summary(hgUnder)

write.csv(Go_under_cluster_15, "Go_under_cluster_15.csv")

```

Analysis of cluster_16

```{r}
hgCutoff <- 0.05

params <- new("GOHyperGParams", geneIds = cluster_16, universeGeneIds = gene_universe,
              annotation = "hgu95av2", ontology = "BP", pvalueCutoff = hgCutoff, conditional = FALSE,
              testDirection = "over")

#Change conditional to true to do a conditional hypogeometric test
hgOver <- hyperGTest(params)
GO_over_cluster_16 <- summary(hgOver)

write.csv(GO_over_cluster_16, "GO_over_cluster_16.csv")


#  test for under represented terms
paramsUnder <- params
testDirection(paramsUnder) <- "under"
hgUnder <- hyperGTest(paramsUnder)
Go_under_cluster_16 <- summary(hgUnder)

write.csv(Go_under_cluster_16, "Go_under_cluster_16.csv")

```

Analysis of cluster_17

```{r}
hgCutoff <- 0.05

params <- new("GOHyperGParams", geneIds = cluster_17, universeGeneIds = gene_universe,
              annotation = "hgu95av2", ontology = "BP", pvalueCutoff = hgCutoff, conditional = FALSE,
              testDirection = "over")

#Change conditional to true to do a conditional hypogeometric test
hgOver <- hyperGTest(params)
GO_over_cluster_17 <- summary(hgOver)

write.csv(GO_over_cluster_17, "GO_over_cluster_17.csv")


#  test for under represented terms
paramsUnder <- params
testDirection(paramsUnder) <- "under"
hgUnder <- hyperGTest(paramsUnder)
Go_under_cluster_17 <- summary(hgUnder)

write.csv(Go_under_cluster_17, "Go_under_cluster_17.csv")

```

Analysis of cluster_18
```{r}
hgCutoff <- 0.05

params <- new("GOHyperGParams", geneIds = cluster_18, universeGeneIds = gene_universe,
              annotation = "hgu95av2", ontology = "BP", pvalueCutoff = hgCutoff, conditional = FALSE,
              testDirection = "over")

#Change conditional to true to do a conditional hypogeometric test
hgOver <- hyperGTest(params)
GO_over_cluster_18 <- summary(hgOver)

write.csv(GO_over_cluster_18, "GO_over_cluster_18.csv")


#  test for under represented terms
paramsUnder <- params
testDirection(paramsUnder) <- "under"
hgUnder <- hyperGTest(paramsUnder)
Go_under_cluster_18 <- summary(hgUnder)

write.csv(Go_under_cluster_18, "Go_under_cluster_18.csv")

```

Analysis of cluster_19
```{r}
hgCutoff <- 0.05

params <- new("GOHyperGParams", geneIds = cluster_19, universeGeneIds = gene_universe,
              annotation = "hgu95av2", ontology = "BP", pvalueCutoff = hgCutoff, conditional = FALSE,
              testDirection = "over")

#Change conditional to true to do a conditional hypogeometric test
hgOver <- hyperGTest(params)
GO_over_cluster_19 <- summary(hgOver)

write.csv(GO_over_cluster_19, "GO_over_cluster_19.csv")


#  test for under represented terms
paramsUnder <- params
testDirection(paramsUnder) <- "under"
hgUnder <- hyperGTest(paramsUnder)
Go_under_cluster_19 <- summary(hgUnder)

write.csv(Go_under_cluster_19, "Go_under_cluster_19.csv")

```

Analysis of cluster_20

```{r}
hgCutoff <- 0.05

params <- new("GOHyperGParams", geneIds = cluster_20, universeGeneIds = gene_universe,
              annotation = "hgu95av2", ontology = "BP", pvalueCutoff = hgCutoff, conditional = FALSE,
              testDirection = "over")

#Change conditional to true to do a conditional hypogeometric test
hgOver <- hyperGTest(params)
GO_over_cluster_20 <- summary(hgOver)

write.csv(GO_over_cluster_20, "GO_over_cluster_20.csv")


#  test for under represented terms
paramsUnder <- params
testDirection(paramsUnder) <- "under"
hgUnder <- hyperGTest(paramsUnder)
Go_under_cluster_20 <- summary(hgUnder)

write.csv(Go_under_cluster_20, "Go_under_cluster_20.csv")

```

Analysis of cluster_21

```{r}
hgCutoff <- 0.05

params <- new("GOHyperGParams", geneIds = cluster_21, universeGeneIds = gene_universe,
              annotation = "hgu95av2", ontology = "BP", pvalueCutoff = hgCutoff, conditional = FALSE,
              testDirection = "over")

#Change conditional to true to do a conditional hypogeometric test
hgOver <- hyperGTest(params)
GO_over_cluster_21 <- summary(hgOver)

write.csv(GO_over_cluster_21, "GO_over_cluster_21.csv")


#  test for under represented terms
paramsUnder <- params
testDirection(paramsUnder) <- "under"
hgUnder <- hyperGTest(paramsUnder)
Go_under_cluster_21 <- summary(hgUnder)

write.csv(Go_under_cluster_21, "Go_under_cluster_21.csv")

```

Analysis of cluster_22
```{r}
hgCutoff <- 0.05

params <- new("GOHyperGParams", geneIds = cluster_22, universeGeneIds = gene_universe,
              annotation = "hgu95av2", ontology = "BP", pvalueCutoff = hgCutoff, conditional = FALSE,
              testDirection = "over")

#Change conditional to true to do a conditional hypogeometric test
hgOver <- hyperGTest(params)
GO_over_cluster_22 <- summary(hgOver)

write.csv(GO_over_cluster_22, "GO_over_cluster_22.csv")


#  test for under represented terms
paramsUnder <- params
testDirection(paramsUnder) <- "under"
hgUnder <- hyperGTest(paramsUnder)
Go_under_cluster_22 <- summary(hgUnder)

write.csv(Go_under_cluster_22, "Go_under_cluster_22.csv")

```


